title: "Propensity Score Matching for Assessing Textile Preferences on Net Use"
author: "Hannah Koenker, Tropical Health"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: cosmo
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = TRUE)
library(tidyverse)
library(janitor)
library(haven)
library(survey)
# install.packages("srvyr")
library(srvyr)
library(labelled)
library(flextable)
library(gtsummary)
library(MatchIt)
theme_set(theme_minimal())
```

# Additional Analyses

## Creating matched groups

We may be able to use propensity score matching approaches to create matched groups of people in PE and PET net-owning households who have similar ‘propensity’ to use ITNs. If there are sufficient numbers of matched individuals, and the residuals are low, we can be reasonably confident that differences in net use would be due to the textile of the net, rather than unmeasured variables. This approach is frequently used in economic and social science research when it is not feasible to randomly assign individuals to experimental groups.

Below, we can use the predictive model from the multivariate regression to calculate the probability of a net having been used the previous night, based on the covariates, and we can compare the distribution of these probabilities across nets that were in fact used, or not used. We are hoping to see that there are bars across a wide range of probabilities for both types, even if the bar heights are not the same.

```{r psm_hist}

# # convert dm to a numeric dataset
# dnum <- lapply(dm, as.numeric(as.character)) 
# # Multivariate
# m2 <- glm(netu ~ textile + supply + hv006 + hv024 + hv025 +hv270 + netagegr, dnum, family = binomial)
# # summary(m1)$coefficients
# tbl_regression(m2, exponentiate = TRUE)
# prs_df2 <- data.frame(pr_score = predict(m2, type = "response"),
#                      used = m2$model$netu)
# df2_nomiss <- dnum %>%  # MatchIt does not allow missing values
#   na.omit()

prs_df <- data.frame(pr_score = predict(m1, type = "response"),
                     used = m1$model$netu)
# head(prs_df)

labs <- paste("Net: ", c("was used", "was not used"))
prs_df %>% 
  mutate(used = ifelse(used==1, labs[1], labs[2])) %>%
  ggplot(aes(x = pr_score,
             fill = used)) +
  geom_histogram(color = "white", bins = 50) +
  facet_wrap(~used) +
  xlab("Probability of net being used the previous night") +
  theme_bw()
```

The next step is to create matched pairs of observations that have the same propensity score (net has the same probability of being used), but differ in whether they were used. 

From the table below we can see that we have produced a dataset that has 2,176 matched pairs of nets - half that were used and half that were not used. Additional information from the covariates and matching process is included in the dataset but not shown below.

```{r psm_match}
netuse_cov <- c("textile",  "supply", "hv006", "hv024", "hv025", "hv270", "netagegr")

df_nomiss <- dm %>%  # MatchIt does not allow missing values
  select(netu, all_of(netuse_cov)) %>%
  na.omit()

# since we have categorical variables, may want to use exact matching (with multiple matches?) and not 'nearest', which is for continuous variables. But this doesn't produce the distance variable.

mod_match <- matchit(netu ~ textile + supply + hv006 + hv024 + hv025 +hv270 + netagegr,
                     method = "nearest", data = df_nomiss)

mod_match2 <- matchit(netu ~ textile + supply + hv006 + hv024 + hv025 +hv270 + netagegr,
                     method = "nearest", data = df2_nomiss)
summary(mod_match)
# plot(mod_match)
plot(mod_match, type = 'jitter', interactive = FALSE)
# plot above shows only a few matched treated units at the high end (.85+) with a lot of matched controls (all of them). Lots of unmatched treated units, below 0.85 propensity score. 

dta_m <- match.data(mod_match)
dim(dta_m) 

dta_m %>% 
  count(netu) %>% 
  flextable()

mod_cem <- matchit(netu ~ textile + supply + hv006 + hv024 + hv025 +hv270 + netagegr,
                     method = "cem", data = df_nomiss)
                   
mod_full <- matchit(netu ~ textile + supply + hv006 + hv024 + hv025 +hv270 + netagegr,
                     method = "full", data = df_nomiss)
plot(mod_full, type = 'jitter', interactive = FALSE)
dta_full <- match.data(mod_full)
dim(dta_full)
dta_full %>% 
  count(netu) %>% 
  flextable()

```

Next, it can be useful to plot the means of the covariates against the estimated propensity score, separating by whether the net was used or not. If the matching was well done, both used and unused nets will have largely identical means of each covariate, for each propensity score value.

```{r psm_visualcheck, include = FALSE}
fn_bal <- function(dta, variable) {
  dta$variable <- dta[, variable]
  dta$variable <- as.numeric(dta$variable)
  dta$netu <- as.factor(dta$netu)
  support <- c(min(dta$variable), max(dta$variable))
  ggplot(dta, aes(x = distance, y = variable, color = netu)) +
    geom_point(alpha = 0.2, size = 1.3) +
    geom_smooth(method = "loess", se = F) +
    xlab("Propensity score") +
    ylab(variable) +
    theme_bw() +
    ylim(support)
}

# dta_n <- lapply(dta_m, as.numeric) 
# support <- c(min(dta_m$hv024), max(dta_m$hv024))
#  geom_smooth(formula = y ~ x, method = "loess", se = F) + 

fn_bal <- function(dta, variable) {
  ggplot(dta, aes(x = distance, y = variable, color = netu)) +
    geom_point(alpha = 0.2, size = 1.3) +
      xlab("Propensity score") +
     theme_bw() 
}

library(gridExtra)
grid.arrange(
   fn_bal(dta_m, "textile"),
   fn_bal(dta_m, "netagegr") + theme(legend.position = "none"),
   fn_bal(dta_m, "hv024"),
   fn_bal(dta_m, "hv025") + theme(legend.position = "none"),
   fn_bal(dta_m, "hv270"),
   fn_bal(dta_m, "hv006") + theme(legend.position = "none"),
   fn_bal(dta_m, "supply"),
   nrow = 4, widths = c(1, 0.8)
)
```

```{r}
dta_m %>% 
  group_by(netu) %>% 
  select(all_of(netuse_cov)) %>% 
  summarise_all(funs(if(is.numeric(.)) mean(., na.rm = TRUE) else first (.)))
  
  #summarise_all(funs(mean))
  
#  summarise_all(funs(if(is.numeric(.)) mean(., na.rm = TRUE) else first (.)))

```

```{r}
lm_treat1 <- lm(textile ~ netu, data = dta_m)
summary(lm_treat1)


```

